---
title: "Workshop"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE, message = FALSE, warning = FALSE)
```

# Hands on!


```{r, message = FALSE}
library(tidyverse)

library(here)
```

#### Import with `read_csv`

```{r, message=FALSE}

got_char <- read_csv(here("01_data", "got_char.csv"))

got_houses <- read_csv(here("______", "________.csv"))

```

## 1. Exercise 1. Top 10 

#### Merge the 2 tables and save it `<-` as a new table named  `got_complete`
```{r}
got_char %>% 
  left_join(got_houses, by = c("actor" = "_____"))
```


#### 2. Make new columns with `mutate`

- Make the the variable `total` as the sum of season 1 to 3, save the updated table

```{r}
got_complete <- ______ %>%
  mutate(total = ________________)
```

- How would you do the average of the three seasons?

```{r}
got_complete <- ______ %>%
  mutate(total = ________________)
```


#### 3. Reorganize your rows with `arrange` 


Ascending
```{r}
got_complete %>% 
  arrange(total)
```


Descending

```{r}
got_complete<- got_complete %>% 
  arrange(desc(_____))
```

#### 4. Select and reorder columns with `select`

```{r}
got_complete %>%
  select(actor, total, house_a) 
```


#### 5. **`rename`** variables

```{r}
got_complete %>% 
  rename(Character = actor,
         House = house_a,
         `Total acting time` = total)
```

#### 6. **`slice`** rows

```{r}
got_complete %>% 
  slice(1:10)
```

#### Pipe all steps and...tada!

```{r}
got_char %>%
  left_join(got_houses, by = c("_____" = "_____")) %>% 
  mutate(_____ = season_1 + season_2 + season_3) %>% 
  arrange(desc(____)) %>% 
  select(actor,____, _____) %>% 
  slice(1:10) %>% 
  rename(Character = ____,
         _____ = house_a,
         `Total acting time` = ____)
```

## Exercise 2. How is the gender distribution across houses?

#### 1. Explore the variables with `count`

How many women and how many men?

```{r}
got_houses %>% 
  count(gender)
```

What is the largest house?
```{r}
got_houses %>% 
  count(house_a, sort = TRUE)
```



#### 2. Drop missing variables with `drop_na`

How many rows do you drop with the following code?

Save this updated file as `got_houses_plot`

```{r}
________________ <- got_houses %>%
  drop_na(gender, house_a)
```

#### 3. Do operations within categories of a variable with `group_by`

**n()** gives the current group size.

```{r}
got_houses_plot <- 
  got_houses_plot %>% 
  group_by(____) %>% 
  mutate(n = n()) %>% 
  ungroup()
```


#### 4. `Filter` rows with a criteria

```{r}
got_houses_plot <- 
  got_houses_plot %>% 
  filter(______) 
```

Try filtering by size and only women

```{r}
got_houses_plot %>% 
  filter(n > ___ & gender == "_____")
```


#### 5. Make labels for gender

**Using `ifelse`, only 2 conditions**

```{r}
got_houses_plot <- got_houses_plot %>%
  mutate(gender = ifelse(gender == 0, "_______", "_________"))
```

```{r}
got_houses_plot %>% 
  count(gender)
```

Tip: Use **`case_when`** for more than two conditions


#### Data ready to be plotted!

```{r}
got_houses_plot <- got_houses %>%
  drop_na(________ , ______) %>%
  group_by(house_a) %>%
  mutate(n = n()) %>%
  ungroup() %>%
  filter(____) %>%
  mutate(gender = ifelse(gender == 0, "_____", "_____"))
```


## `ggplot` grammar

#### 1. Let's start with a basic bar plot


```{r}
got_houses_plot %>%
  ggplot(aes(____)) +
  geom_bar()
```
 
#### 2. Now let's add add gender to aes()


```{r }
got_houses_plot %>%
  ggplot(aes(_____, fill = ____)) +
  geom_bar()
```

#### 3. Flip the coords


```{r}
got_houses_plot %>%
  ggplot(aes(__ = house_a, 
             fill= gender)) +
  geom_bar()
```

Before we needed to use `coord_flip()`

#### 4. Sort by frequency


```{r}
got_houses_plot %>%
  ggplot(aes(y = reorder(house_a, ___),
             fill= gender)) +
  geom_bar()
```

#### 5. Details count &#x1F485;&#x1F3FD;

```{r fig.width = 12}
got_houses_plot %>%
  ggplot(aes(y = reorder(house_a, __), fill= gender)) +
  geom_bar() +
  labs(title = "Distribution of gender across the houses",
       x = "Number of characters",
       y = "House",
       fill = "Gender") +
  theme_minimal()
```

## Exercise 3. How was the evolution of the protagonists across seasons?

#### Go from wide to long with `pivot_longer`

Save the new long data as `got_long`
```{r}
_____ <- got_complete %>%
  pivot_longer(
    cols = season_1:season_7,
    names_to = "_____",
    values_to = "_____",
    names_prefix = "season_")

head(got_long)

```

#### Create a total sum of time by character

**`group_by() + summarise`**

```{r, message = FALSE}
got_long %>%
  group_by(actor) %>%
  summarise(total = sum(____)) %>%
  ungroup () 
```

**`group_by() + mutate`**

```{r}
got_long <- got_long %>%
  group_by(actor) %>%
  mutate(total = sum(____)) %>%
  ungroup()
```

### Back to the graph

#### 1. Add the aesthetics and geoms


```{r}
got_long %>%
  ggplot(aes(season, time))+
  geom_point() +
  geom_line()
```

#### 2. Add actors to aes()


```{r}
got_long %>%
  ggplot(aes(season, time, group = _____)) +
  geom_point() +
  geom_line()
  
```

#### 3. Filter the top ten (total > 130min)


```{r}
got_long %>%
  filter(________) %>%
  ggplot(aes(season, time, group = _____)) +
  geom_point() +
  geom_line()
```


#### 4. Add a color for each character


```{r}
got_long %>%
  filter(_____) %>%
  ggplot(aes(season, time, group = _____, color = _____)) +
  geom_point() +
  geom_line()
```

#### 5. Details

```{r}
got_long %>%
  filter(_____) %>%
  ggplot(aes(season, time, group = _____, color = _____)) +
  geom_point() +
  geom_line() +
  theme_minimal() +
  labs(title = "Evolution of the protagonists across seasons",
       x = "Season",
       y = "Total time (min)",
       color = "Protagonist") +
  theme_minimal()
```
